The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-04-01 Paper/jaccardbaetica.tex
1st draft

\maria{deberia cerrar esto}

@@

The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
millet_anforas_1998}. This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}. Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown. Indeed most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.

@@

Frequently, they were marked mainly in handles, but seldom in rims and body \citep{millet_anforas_1998}. 
The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion. Stamps are mostly formed by a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}. Despite researchers identify stamps as identity marks, there is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}. Alike most Dressel 20 containers were not stamped being difficult to determinate.

@@

Our case study examines the relationship between the distribution of amphoric stamps and the workshops. In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps. Based on this assumption, we proposed three hypotheses: a) it exists a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.

@@

It is well known that this product was not frequently consumed by indigenous population. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman legions.

At this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[161]{monfort_britanniaen_1998}. This fact will be spatial incidence in sites close at Hadrian Wall.  
Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}. The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges. Thus, the network was mainly focused on the support of soldiers during the military campaigns. 
%debate intercambios puntuales o redes comerciales de abastecimiento

@@

The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto. The advance of the Roman legions toward \textit{Gallia} opened a way through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
remesal_baetica_2002}.

@@

The analysis shows that amphoric stamps could be correlated with spatial distance. The correlation coefficients range from a minimum to a maximum. The dendrogram in Fig. \ref{dendro} was obtained with Morisita-Horn index. The dendrogram suggests that amphora workshops used different stamps for their production system. Nearby workshops show similarity on the stamps while most of them seem to display different stamps roles. Additionally, groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed of no more than three workshops. Indeed, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.

@@

\section{Conclusion remarks}

@@

On the other hand,the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres. This could be interpreted such as batch systematic organisation between potters \citep{juanmorostesis}. 
This method allowed to potters organise the production with the batch stamping for the posterior commercialisation.

>>>>

@@

The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
millet_anforas_1998}. This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}. Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown. Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.

@@

Frequently, they were mainly marked  in handles, but seldom in rims and body \citep{millet_anforas_1998}. 
The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion. Stamps are mostly formed by a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}. Despite researchers identify stamps as identity marks, there is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}. Alike most Dressel 20 containers were not stamped being difficult to determinate.

@@

Our case study examines the relationship between the distribution of amphoric stamps and the workshops. In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps. Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.

@@

It is well known that this product was not frequently consumed by indigenous population. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.

At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}. This fact will be spatial incidence in sites close at Hadrian Wall.  
Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}. The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges. Thus, the network was mainly focused on the support of soldiers during the military campaigns. 
%debate intercambios puntuales o redes comerciales de abastecimiento

@@

The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto. The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
remesal_baetica_2002}.

@@

The analysis shows that amphoric stamps could be correlated with spatial distance. The correlation coefficients range from a minimum to a maximum. The dendrogram in Fig. \ref{dendro} was obtained with Morisita-Horn index. The dendrogram suggests that amphora workshops used different stamps for their production system. Nearby workshops show similarity on the stamps while most of them seem to display different stamps roles. Groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed of no more than three workshops. Additionally, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.

@@

\section{Conclusion}

@@

On the other hand, the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres. This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}. 
This method allowed to potters organise the production with the batch stamping for the posterior commercialisation.
